Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-03-26 introduction.tex
introduction - some pruning.

% Historically, the ATLAS Computing model~\cite{jones2008atlas} is based on a
% Grid paradigm~\cite{foster2003grid}, with multilevel, hierarchically
% distributed computing and storage resources.

% PanDA has been developed to meet
% growing ATLAS production and analysis requirements for a data-driven workload
% management system capable of operating at LHC data processing scale.

% PanDA has a highly scalable architecture. Scalability has been demonstrated in
% ATLAS through the rapid increase in usage over the past several years of
% operations, and PanDA is expected to meet the continuously growing computing
% requirements of ATLAS over the next decade. PanDA was designed to have the
% flexibility to adapt to emerging computing technologies in processing, storage,
% networking as well as the underlying software stack. This flexibility has also
% been successfully demonstrated through the past six years: computing centers in
% ATLAS, spanning many continents, were seamlessly integrated into PanDA\@.

% PanDA manages a wide spectrum of workloads, ranging from raw data processing to
% Monte Carlo simulation and user analysis, while constantly evolving to meet
% rapidly changing science needs.

\mtnote{Used the following in related work. We may want to reduce/remove it from
the introduction: ``Today, PanDA serves several thousand users, managing job
distribution to hundreds of ATLAS sites with more than 100,000 CPU cores which
process more than a million jobs per day.''} % (Figure~\ref{fig:daily}).

% \begin{figure}
%     \begin{center}
%         \includegraphics[width=\columnwidth]{figures/DailyJobs.png}
%         \caption{Daily completed jobs on ATLAS Grid for the past 12 month}
%     \end{center}
% \label{fig:daily}
% \end{figure}

In this paper, we describe how PanDA has been engineered to execute a specific
stage of the ATLAS Monte Carlo workflow on Titan, the larger high-performance
computing HPC system currently available in the USA\@.\mtnote{Explain the
benefits offered by Titan in terms of multithreading per node and possibly large
amount of concurrent nodes. Introduce also the notion of backfill.} This extends
the scope of PanDA's compute model, integrating both high-throughput and
high-performance computing resources and enabling the concurrent execution of
both  single and multi-core jobs. The integration of PanDA and Titan went
through three main engineering phases: (i) feasibility study and rapid
prototyping of an initial solution; (ii) progressing scaling of the  prototype
to saturate the available resources; (iii) study of a product-grade architecture
for generic HPC resources. Both phase i and ii have been completed enabling the
execution of up to eight million jobs a week on Titan. A prototype has been
engineered to support phase iii and experimental data are being collected.

In the next section we introduce \ldots.

Why Titan?
\begin{itemize}
    \item A lot of slow (relative the grid) and homogeneous cores together
    \item Grid is saturated
    \item Enable HPC as a calss of computing resource
    \item Enable future DOE experimental and observational capabilities on HPC
    \item Why is so important: more data, run 2 and run 3.
    \item growth of grid is flat (economic model) and saturated. We need more CPU because we have more data.
    \item ATLAS spend most of the time on simulations this is why we want to offload simulations - that happen happen to be performed via AthenaMP.
\end{itemize}

\subsubsection*{The following is a candidate for major compression into a single paragraph}

\mtnote{Moved from related work were we use already this terminology. To be
iterated/adjusted for consistency once the first draft will be ready.}

The term ``workflow'' is used in many disciplines with different meaning. In the
field of scientific computing, ``workflow'' assumes different meanings depending
on the characteristics of the computation, of the software tools used to support
this computation, and of the resources on which it is performed. Further, a
workflow may indicate a whole application, a description of the computational
process of that application or, more commonly, a series of tasks related by data
dependences.

The lack of a consistent and shared definition of ``workflow'' hinders the
understanding of its properties and its relations with related concepts. For
example, we need to clarify the difference among ``workflow'', ``workload'',
``task'', or ``job'' but also between workflow ``template'' and ``instance'', or
``data-flow'' and ``control-flow''. This is precondition to specify properly the
design of software systems that support the execution of scientific
applications.

In this paper, we use the following definitions:

\begin{description}

  \item[Task.] A set of operations to be performed on a computing platform,
  alongside a description of the properties and dependences of those operations,
  and indications on how the dependences should be satisfied and the operations
  should be executed.

  \item[Job.] A unit of  work performed by submitting a script to a resource
  management system (LRMS), like  Slurm or PBS, or by requesting a virtual
  machine or a container to a site supporting virtualization. One or more jobs
  can perform the operations described with a task.

  \item[Workload.] A set of jobs that can be executed concurrently, possibly
  related by a set of relations. For example, jobs of a workload can share one
  or more input files or communicate during execution.

  \item[Workflow.] Set of jobs, related by a set of relations that define the
  order in which each task can be executed. Data dependences are the most common
  relations among workloads, used to define the precedence among their
  executions.

\end{description}

>>>>

% \subsubsection*{The following is a candidate for major compression into a single paragraph}

% \mtnote{Used the following in related work. We may want to reduce/remove it from
% the introduction: ``Today, PanDA serves several thousand users, managing job
% distribution to hundreds of ATLAS sites with more than 100,000 CPU cores which
% process more than a million jobs per day.''} % (Figure~\ref{fig:daily}).

% In this paper, we describe how PanDA has been engineered to execute a specific
% stage of the ATLAS Monte Carlo workflow on Titan, the larger high-performance
% computing HPC system currently available in the USA\@.\mtnote{Explain the
% benefits offered by Titan in terms of multithreading per node and possibly large
% amount of concurrent nodes. Introduce also the notion of backfill.} This extends
% the scope of PanDA's compute model, integrating both high-throughput and
% high-performance computing resources and enabling the concurrent execution of
% both  single and multi-core jobs. The integration of PanDA and Titan went
% through three main engineering phases: (i) feasibility study and rapid
% prototyping of an initial solution; (ii) progressing scaling of the  prototype
% to saturate the available resources; (iii) study of a product-grade architecture
% for generic HPC resources. Both phase i and ii have been completed enabling the
% execution of up to eight million jobs a week on Titan. A prototype has been
% engineered to support phase iii and experimental data are being collected.

% In the next section we introduce \ldots.

% Why Titan?
% \begin{itemize}
%     \item A lot of slow (relative the grid) and homogeneous cores together
%     \item Grid is saturated
%     \item Enable HPC as a calss of computing resource
%     \item Enable future DOE experimental and observational capabilities on HPC
%     \item Why is so important: more data, run 2 and run 3.
%     \item growth of grid is flat (economic model) and saturated. We need more CPU because we have more data.
%     \item ATLAS spend most of the time on simulations this is why we want to offload simulations - that happen happen to be performed via AthenaMP.
% \end{itemize}

% \mtnote{Moved from related work were we use already this terminology. To be
% iterated/adjusted for consistency once the first draft will be ready.}

% The term ``workflow'' is used in many disciplines with different meaning. In the
% field of scientific computing, ``workflow'' assumes different meanings depending
% on the characteristics of the computation, of the software tools used to support
% this computation, and of the resources on which it is performed. Further, a
% workflow may indicate a whole application, a description of the computational
% process of that application or, more commonly, a series of tasks related by data
% dependences.

% The lack of a consistent and shared definition of ``workflow'' hinders the
% understanding of its properties and its relations with related concepts. For
% example, we need to clarify the difference among ``workflow'', ``workload'',
% ``task'', or ``job'' but also between workflow ``template'' and ``instance'', or
% ``data-flow'' and ``control-flow''. This is precondition to specify properly the
% design of software systems that support the execution of scientific
% applications.

% In this paper, we use the following definitions:

% \begin{description}

%   \item[Task.] A set of operations to be performed on a computing platform,
%   alongside a description of the properties and dependences of those operations,
%   and indications on how the dependences should be satisfied and the operations
%   should be executed.

%   \item[Job.] A unit of  work performed by submitting a script to a resource
%   management system (LRMS), like  Slurm or PBS, or by requesting a virtual
%   machine or a container to a site supporting virtualization. One or more jobs
%   can perform the operations described with a task.

%   \item[Workload.] A set of jobs that can be executed concurrently, possibly
%   related by a set of relations. For example, jobs of a workload can share one
%   or more input files or communicate during execution.

%   \item[Workflow.] Set of jobs, related by a set of relations that define the
%   order in which each task can be executed. Data dependences are the most common
%   relations among workloads, used to define the precedence among their
%   executions.

% \end{description}
